% Co-governed and enforced under the Sovereign Integrity Protocol License (SIP License v1.1):
% https://github.com/tensorrent/ACS-Framework/blob/main/LICENSE
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts,amsthm,bm}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{xcolor}
\usepackage{float}
\usepackage{enumitem}
\usepackage{geometry}
\geometry{margin=0.75in}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\hypersetup{
    colorlinks=true,
    linkcolor=blue!70!black,
    citecolor=green!50!black,
    urlcolor=red!70!black
}

\newcommand{\dd}{\mathrm{d}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\ee}{\mathrm{e}}

\title{\LARGE\bfseries Density Engine and Many-Worlds Self-Similarity:\\
Geometric Constraint Routing in the Flag Condensate Programme}

\author{\textbf{ACS Theoretical Physics Working Group}\\[0.3em]
\small\textit{Sovereign-Stack ACS Research Program \& Flag Condensate Project}\\
\small\texttt{research@acs-foundation.org}}

\date{July 23, 2026}

\begin{document}

\maketitle

\begin{abstract}
Two linked interpretive constructs are formalised under RC1 claim discipline.
\textbf{(1) Density engine:} the shape of the universe is read, in this ontology,
as an \emph{inexhaustible geometric constraint source}---geometry routes,
compresses, and re-releases phase structure along lanes; ``density'' means
structure per geometric lane on a stated manifold, not unqualified infinite
J/m$^3$.  \textbf{(2) Many-worlds self-similarity:} the same phase-slip /
throat / eigen-path pattern is postulated to recur across electron, nuclear,
and cosmological scales as structural recursion, not as a proof of Everett's
many-worlds interpretation or as uniquely derived physics.  Popular MWI/string
``every plausibility in parallel'' is \textbf{not} literal ontological world
multiplication here: it reads as \emph{polymorphic quantum flags}---one
$\Phi$ substrate, many phase/mode configurations, \textbf{Lisp-deterministic}
(not parallel-world stochastic): counterfactuals as opcodes on a fixed
substrate, not ontological multiplication.  \textbf{Nested determinants:}
one playthrough substrate with a stack of conditional flags (RPG quest/karma/rep
\emph{language} only); hierarchical guard predicates gate opcode paths; macro
expansion at scale $N$ wraps micro opcodes.  Formal numerics
for electron geometry, nuclear decay, and Hawking recovery remain scoped to
companion notes~\cite{Mobius2026,FlagDecay2026,PalphaRefined2026}; the Klein-Foam
Monad supplies broader ontological framing~\cite{KFM2026}.  A short DAW/session
metaphor appendix links the eigen-path visualisation harness to polymorphic
mode-lane language on one session without enlarging the theorem surface.
\end{abstract}

\vspace{0.1em}
\hrule
\vspace{0.4em}

\section{Introduction and RC1 scope}
\label{sec:intro}

This note is an \textbf{interpretive companion} to the Flag Condensate
programme.  It separates three layers:

\begin{enumerate}[leftmargin=1.5em,itemsep=2pt]
  \item \textbf{Model claims} --- finite statements on stated manifolds with
        cited numerics (electron capacitance, Geiger--Nuttall slope,
        transfer-matrix Bogoliubov identity, $P_\alpha$ overlap proxies).
  \item \textbf{Structural parallels} --- the same phase-defect template
        applied across domains in Table~\ref{tab:unification}; identity beyond
        structural form is not claimed.
  \item \textbf{Metaphor / sovereign framing} --- density engine, Monad
        recursion, DAW branching lanes; narrative energy without physics proof.
\end{enumerate}

\begin{remark}[RC1 discipline]
``Infinite density engine'' denotes an \emph{inexhaustible geometric constraint
source}: no finite routing exhausts the phase-slip supply on the postulated
foam.  It does \textbf{not} assert literal infinite energy density in SI units
unless a finite manifold, regularisation, and measurement protocol are stated.
``Many worlds self-similar'' denotes \emph{scale recursion of the same
phase-defect pattern}; it is not a claim that MWI is established physics or
uniquely derived from this programme.  ``Every plausibility in parallel'' (popular
MWI/string gloss) is \textbf{not} literal ontological world multiplication in
flag ontology: it denotes \emph{polymorphic quantum flags} on one $\Phi$
substrate---Lisp-deterministic conditional structure (opcode table on
$\Phi$), not parallel branches all ``running''---an alternative reading,
not a disproof of MWI or string theory.
\end{remark}

\section{The density engine (scoped)}
\label{sec:density}

\subsection{Definition}

\begin{definition}[Geometric lane density]
On a stated throat or routing manifold $\mathcal{M}$ with coordinate $r$ and
effective metric $g_{ab}$, define \emph{lane density} as phase-structure
content per unit geometric route:
\begin{equation}
\rho_{\mathrm{lane}}(r) \equiv
\frac{|\Phi(r)|^2}{V_{\mathrm{eff}}(r)}\,,
\end{equation}
where $V_{\mathrm{eff}}$ is the effective volume element of the lane and
$\Phi=(\phi+\ii\pi)/\sqrt{2}$ is the flag condensate amplitude.  On the
nuclear throat, $\rho_{\mathrm{lane}}$ tracks confined standing-wave content
before spectral bifurcation~\cite{FlagDecay2026}; on the electron annulus,
capacitance routing reads stored twist energy per geometric turn~\cite{Mobius2026}.
\end{definition}

\subsection{Engine metaphor vs.\ model}

In \textbf{metaphor}: the universe-as-shape acts as a density \emph{engine}---
geometry is not passive container but active router.  Hypercone throats compress
phase; phase slips re-release structure along new lanes; slag deposits as mass
read in the Klein-Foam ontology~\cite{KFM2026}.

In \textbf{model} (stated scope only): $\rho_{\mathrm{lane}}$ is finite on
each cited manifold.  The nuclear decay note reports Gamow action
$W=\delta S(R)$ and Bogoliubov ratio $|\beta/\alpha|^2=\ee^{-2W}$ on 14
alpha emitters~\cite{FlagDecay2026}.  The refined $P_\alpha$ overlap note
reports correlation $r\approx 0.89$ between geometric overlap proxy and
extracted preformation on the same set~\cite{PalphaRefined2026}.  No claim is
made that $\rho_{\mathrm{lane}}\to\infty$ pointwise in physical units.

\subsection{Routing operations}

Three engine operations recur in the programme lexicon:

\begin{itemize}[leftmargin=1.3em,itemsep=2pt]
  \item \textbf{Route} --- direct $\phi$/$\pi$ quadrature along a throat
        geodesic (confined $j_\ell$ modes inside $r<R$).
  \item \textbf{Compress} --- accumulate phase slip across colour-confining
        curvature (instanton deformation $\delta S$).
  \item \textbf{Re-release} --- spectral bifurcation to travelling wave
        (alpha packet, Hawking mode mix, fluxon nucleation).
\end{itemize}

These are \emph{readings} of the transfer-matrix spine in~\cite{FlagDecay2026},
not independent postulates.

\section{Self-similar micro--macro structure}
\label{sec:selfsim}

\subsection{Scale ladder}

The programme postulates that one phase-defect template recurs across scales:

\begin{table}[H]
\centering
\caption{Self-similar phase-defect ladder (interpretive; numerics scoped per row).}
\label{tab:ladder}
\small
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Scale} & \textbf{Geometry read} & \textbf{Formal anchor} \\
\midrule
Micro & Framed unknot electron; $Sl=2$ & \cite{Mobius2026} \\
Meso & Nuclear colour throat; Gamow $W$ & \cite{FlagDecay2026,PalphaRefined2026} \\
Macro & Schwarzschild horizon; Hawking $T_H$ & \cite{FlagDecay2026,KFM2026} \\
\bottomrule
\end{tabular}
\end{table}

At each rung: confined standing wave $\to$ phase slip at defect $\to$
Bogoliubov mode mixing $\to$ observable ( $g$-factor proxy, half-life,
Hawking temperature).  \textbf{Self-similarity} means structural recurrence of
this \emph{pattern}, not identity of numerical coefficients or a single
universal action without domain-specific boundary data.

\subsection{Monad recursion}

The Klein-Foam Monad~\cite{KFM2026} supplies ontological language: local
structures are reflections of whole foam self-intersection.  In RC1 terms this
is \textbf{sovereign framing}---useful for programme coherence, not a
substitute for domain-specific verification.

\section{Many-worlds as structural parallel}
\label{sec:mwi}

\subsection{Polymorphic quantum flags (core correction)}

Popular MWI/string language---``every plausibility in parallel''---is
\textbf{not} read as literal ontological world multiplication in this flag
ontology.  The programme instead describes \textbf{polymorphic quantum flags}:
one $\Phi$ substrate supporting many phase/mode configurations after spectral
bifurcation.  Branching is \emph{mode polymorphism on one session}, analogous
to polymorphic DAW lanes, not a census of multiplied worlds.

\textbf{Lisp-deterministic read (RC1).}  Polymorphic flags are
\textbf{Lisp-deterministic}, not parallel-world stochastic.  Counterfactuals
read as: \emph{this would have happened / could have happened if you had done
$x$}---like \textbf{opcodes} on a fixed substrate.  Conditional structure is
encoded in the flag program, not ontological multiplication.  One deterministic
evaluation context; admissible morphs are conditional paths
(\texttt{if}/\texttt{when}/\texttt{unless} on phase geometry)---an
\textbf{opcode table on $\Phi$}.

\textbf{Nested determinants (sovereign framing).}  Branching RPG systems
(e.g.\ Fable karma tiers, Fallout faction reputation, quest flags) model
\emph{nested determinants} on one playthrough substrate---not parallel save
files all running.  Quest state, karma tier, and faction rep read as nested
determinant layers on $\Phi$; branch dialogue is opcode table lookup gated
by outer predicates, not a live forked universe.  \emph{Would/could if $x$}
denotes nested predicate evaluation at each layer; deterministic given the
guard stack; outer layers gate inner opcode paths.

\textbf{Formal read (scoped).}  \textbf{Hierarchical guard predicates} on
phase-slip paths.  Macro expansion at scale $N$ (meso/macro ladder) wraps
micro opcodes; self-similar micro--macro recursion is the same nesting
pattern at different resolution.  DAW compiled automation curves are
pre-expanded opcode sequences on one timeline.

\begin{remark}[RC1: game analogies]
RPG/quest-flag language is \textbf{interpretive sovereign framing}.  It
does \textbf{not} assert that game mechanics prove physics, nor disprove
MWI or string theory.
\end{remark}

\begin{table}[H]
\centering
\caption{Misread vs.\ model read (RC1: alternative reading, not disproof of MWI/string).}
\label{tab:misread}
\small
\begin{tabular}{@{}p{0.42\linewidth}p{0.48\linewidth}@{}}
\toprule
\textbf{Misread (popular MWI/string gloss)} & \textbf{Model read (flag ontology)} \\
\midrule
Infinite parallel branches all ``running'' &
One deterministic evaluation context on $\Phi$ \\
Infinite parallel playthroughs all ``running'' &
One session, \textbf{nested determinant stack} on $\Phi$ \\
Every plausibility exists as a separate ontological world &
One $\Phi$ substrate; admissible morphs = conditional paths (opcode table) \\
Parallel worlds multiply physically &
Polymorphic mode lanes on one session---compiled opcodes, not parallel runs \\
Quest branches = forked save universes &
Quest/karma/rep flags = nested guard predicates; dialogue = opcode lookup \\
``Could have'' = live alternate universe &
``Could have'' = nested predicate on guard stack (Lisp macro expansion) \\
Branch count is metaphysical inventory &
Branch language = spectral bifurcation output, not world census \\
String landscape = literal parallel universes \emph{here} &
Structural parallel only; scoped alternative reading \\
Everett branching resolved in favour of this programme &
RC1: alternative reading, not disproof of MWI/string \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Lisp and opcode analogies (sovereign framing)}

\begin{table}[H]
\centering
\caption{Structural analogies for polymorphic flags (metaphor; not literal interpreter claim).}
\label{tab:opcode}
\small
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Analogy} & \textbf{Flag-programme read} \\
\midrule
Lisp macros & Expand to deterministic s-expressions; counterfactual =
unevaluated branch in the form, not a live universe \\
Quest flag / karma tier & Outer \textbf{guard predicate} on opcode table
(nested determinant layer) \\
Faction reputation & Nested determinant; gates which dialogue opcodes are
admissible \\
Branch dialogue & Opcode table lookup on $\Phi$, not live forked universe \\
Phase slip & \texttt{jump}/\texttt{call} opcode \\
Throat & Branch target (label / entry point) \\
Gamow envelope & Guard predicate (\texttt{when} on phase geometry) \\
$S(A,Z)$ & Per-isotope immediate operand (spectroscopic clip) \\
Scale ladder (micro$\to$macro) & Macro expansion at $N$ wrapping micro opcodes \\
DAW automation curves & Compiled opcodes on the timeline---not parallel sessions \\
\bottomrule
\end{tabular}
\end{table}

\begin{remark}[RC1 boundary]
Lisp/opcode/game language is \textbf{sovereign framing} and structural metaphor.
It does not assert a literal Lisp interpreter or RPG engine in nature, nor that
games prove physics, nor disprove MWI or string theory.
\end{remark}

\subsection{What is \emph{not} claimed}

\begin{itemize}[leftmargin=1.3em,itemsep=2pt]
  \item Everett's many-worlds interpretation (MWI) is \textbf{not} proven or
        uniquely derived here.
  \item Branching is \textbf{not} asserted as empirical fact at all scales.
  \item Ontological world multiplication is \textbf{not} claimed (no literal
        parallel-universe inventory).
  \item Parallel worlds are \textbf{not} counted or predicted numerically in
        this note.
  \item MWI or string theory is \textbf{not} disproved---only an alternative
        reading scoped to flag ontology is offered.
\end{itemize}

\subsection{What \emph{is} claimed}

\textbf{Structural parallel:} when spectral bifurcation splits a confined mode
into outgoing components, the transfer matrix naturally produces multiple
consistent continuations---\emph{polymorphic quantum flags} on one $\Phi$
substrate, analogous to polymorphic mode lanes on one session or parallel
automation lanes in the eigen-path DAW visualisation (Appendix~\ref{app:daw}).
In bubble-cascade collapse language~\cite{KFM2026}, local reconfiguration can
be \emph{read} as many consistent foam patches without asserting MWI as
established physics or ontological world multiplication.

\begin{remark}
RC1: treat ``many worlds'' here as \emph{polymorphic phase/mode configurations
on one $\Phi$ substrate under the same geometric template}---Lisp-deterministic
conditional paths (opcode table), nested determinant guard stack on one session,
not parallel branches all running, not ontological world multiplication, and not
a claim that measurement or decoherence
is resolved in favour of Everett over Copenhagen, many-minds, or other
interpretations.
\end{remark}

\section{Link to four-domain unification table}
\label{sec:unification}

Table~\ref{tab:unification} reproduces the phase-defect unification row set
from~\cite{FlagDecay2026}.  The density engine reads each row as a different
\emph{lane} on the same Bogoliubov router; many-worlds self-similarity reads
the table as the same pattern at different scale/boundary limits.

\begin{table}[H]
\centering
\caption{Four-domain phase-defect unification (structural parallel only).}
\label{tab:unification}
\small
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Domain} & \textbf{Phase defect $W$} & \textbf{Observable} \\
\midrule
Nuclear decay ($^{238}\mathrm{U}$) & 41.517 & $t_{1/2}=4.47\times 10^9$\,yr \\
Black hole ($1\,M_\odot$) & 0.000 & $T_H=6.17\times 10^{-8}$\,K \\
Sphaleron (electroweak) & 90.000 & $B+L$ violation rate \\
Superconductor (NbTi) & 17.400 & Fluxon nucleation rate \\
\bottomrule
\end{tabular}
\end{table}

Absolute physical identity beyond shared transfer-matrix \emph{form} is not
claimed.  Hawking recovery uses horizon surface gravity as boundary data;
nuclear decay uses Coulomb throat data; sphaleron and fluxon rows are
programme placeholders at stated $W$~\cite{FlagDecay2026}.

\section{Conclusion}
\label{sec:conclusion}

The density engine names geometry-as-router: finite lane density on stated
manifolds in model mode; inexhaustible constraint supply in metaphor mode.
Many-worlds self-similarity names scale recursion of phase-slip / throat /
eigen-path structure across electron, nuclear, and cosmological readings---a
structural parallel naming polymorphic flags on one substrate, not MWI proof or
ontological world multiplication.  All quantitative claims remain scoped to
\cite{Mobius2026,FlagDecay2026,PalphaRefined2026,KFM2026}.

\appendix
\section{DAW / session metaphor (non-formal)}
\label{app:daw}

The eigen-path DAW visualisation (\texttt{Aiso\_build\_artifacts/eigen\_path\_daw\_viz/})
renders radial profiles as mixer lanes: $u_{\mathrm{in}}$, $u_\alpha$, throat
weight, overlap bus, Gamow envelope, spectroscopic clips.  \textbf{Metaphor
map:}

\begin{itemize}[leftmargin=1.3em,itemsep=2pt]
  \item \textbf{Timeline} $r$ (fm) $\leftrightarrow$ routing coordinate along
        throat lane.
  \item \textbf{Polymorphic mode lanes} $\leftrightarrow$ coexisting
        phase/mode configurations on one $\Phi$ substrate after bifurcation
        (many-worlds \emph{language}, not ontological world multiplication).
  \item \textbf{Automation curves} $\leftrightarrow$ \emph{compiled opcodes}
        on the timeline---conditional envelopes and clips, not parallel DAW
        sessions (Lisp-deterministic read).
  \item \textbf{Opcode map} (Table~\ref{tab:opcode}): phase slip = jump/call;
        throat = branch target; Gamow envelope = guard predicate;
        $S(A,Z)$ = per-isotope immediate operand.
  \item \textbf{Global S fader} $\leftrightarrow$ cross-lane gain matching
        extracted $P_\alpha$ intercept~\cite{PalphaRefined2026}.
  \item \textbf{Nested determinant stack} $\leftrightarrow$ quest flags /
        karma tiers / faction rep (RPG \emph{language} only): outer guards
        gate inner opcode paths on one $\Phi$ session.
  \item \textbf{Scale ladder canvas} (\texttt{density-engine.canvas.tsx})
        $\leftrightarrow$ micro / meso / macro rungs of
        Table~\ref{tab:ladder}.
\end{itemize}

This appendix does not enlarge the theorem or measurement surface of any
companion note.

\begin{thebibliography}{99}
\bibitem{Mobius2026}
Flag Condensate Collaboration,
\emph{The M\"obius-Screw Electron: Framed Unknot Geometry for $g=2$ and a Capacitance Model of $\alpha$},
ACS Framework note (July 2026),
\texttt{papers/notes/Mobius\_Screw\_Electron.tex}.

\bibitem{FlagDecay2026}
Flag Condensate Collaboration,
\emph{Nuclear Decay as a Spectral Bifurcation of the Flag Condensate},
ACS Framework note (July 2026),
\texttt{papers/notes/Flag\_Condensate\_Nuclear\_Decay.tex}.

\bibitem{KFM2026}
B. Wallace,
\emph{The Klein-Foam Monad: A Geometric Postulation of Reality},
ACS Framework note (July 2026),
\texttt{papers/notes/Klein\_Foam\_Monad.tex}.

\bibitem{PalphaRefined2026}
ACS Theoretical Physics Working Group,
\emph{Refined $P_\alpha$ Overlap: Parametric Spectroscopic Factor and Self-Consistent WS+Coulomb Eigenmode},
ACS Framework note (July 2026),
\texttt{flag\_condensate\_palpha\_refined.tex} /
\texttt{papers/notes/Flag\_Condensate\_Palpha\_Refined.tex}.

\bibitem{PalphaOverlap2026}
ACS Theoretical Physics Working Group,
\emph{Standing-Wave Overlap Proxy for Alpha Preformation},
ACS Framework note (July 2026),
\texttt{flag\_condensate\_palpha\_overlap.tex}.
\end{thebibliography}

\end{document}
