% Co-governed and enforced under the Sovereign Integrity Protocol License (SIP License v1.1):
% https://github.com/tensorrent/ACS-Framework/blob/main/LICENSE
\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts,amsthm,bm}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{xcolor}
\usepackage{float}
\usepackage{geometry}
\geometry{margin=0.60in}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\hypersetup{
    colorlinks=true,
    linkcolor=blue!70!black,
    citecolor=green!50!black,
    urlcolor=red!70!black
}

\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
\setlength{\abovedisplayshortskip}{1pt}
\setlength{\belowdisplayshortskip}{1pt}

\newcommand{\dd}{\mathrm{d}}
\newcommand{\ee}{\mathrm{e}}

\title{\LARGE\bfseries Throat-Geometry Overlap for Alpha Preformation:\\
Warped Measure and Woods--Saxon Cluster Trial}

\author{\textbf{ACS Theoretical Physics Working Group}\\[0.3em]
\small\textit{Sovereign-Stack ACS Research Program \& Flag Condensate Project}\\
\small\texttt{research@acs-foundation.org}}

\date{July 22, 2026}

\begin{document}

\maketitle

\begin{abstract}
The flat $[0,R]$ standing-wave overlap proxy~\cite{PalphaOverlap2026}
uses $\ell=0$ flag mode vs Gaussian alpha trial with measure $\dd r$.
This sequel replaces that proxy geometry with (i)~the AdS-like throat
radial weight $w(r)=R/r$ implied by the companion nuclear metric
$\dd s^2=(R^2/r^2)\,\dd r^2+\cdots$~\cite{NuclearDecay2026}, and
(ii)~a Woods--Saxon (WS) alpha-cluster radial trial as primary, keeping
Gaussian/HO as comparison.  Overlap support remains the confined interior
$[0,R]$ where the Dirichlet flag mode is defined; the barrier $[R,b]$ stays
the Gamow domain.  On the same 14-isotope set, primary throat+WS yields
$\langle\log_{10} P_{\mathrm{model}}\rangle=-0.4607$, Pearson
$r=+0.8875$ vs extracted $\log_{10} P_\alpha$, and
$S_\star\approx 3.18\times 10^{-2}$ reducing RMS residual from $1.7099$
to $0.8247$.  Relative to the flat-Gaussian proxy, correlation changes by
$\Delta r\approx -0.0007$; absolute-scale $S$ shifts modestly.  Still a
geometric/structural proxy---not microscopic uniqueness.
\end{abstract}

\vspace{0.1em}
\hrule
\vspace{0.4em}

\section{Motivation and what moved}

Reference~\cite{PalphaOverlap2026} documented a flat-measure overlap
proxy and flagged full throat geometry and a more realistic alpha trial
as the next structural step.  This note implements that step on the
unchanged 14-isotope table of~\cite{NuclearDecay2026}.

\begin{remark}[RC1 scope]
$P_{\mathrm{model}}$ remains a geometric/structural preformation
\emph{proxy}.  Warped measure and WS shape are mechanism choices
documented below; they are not a claim of unique many-body nuclear
structure.  Absolute scale may still need one global factor $S$.
\end{remark}

\subsection{What moved (proxy $\to$ throat/WS)}
\begin{itemize}
\item \textbf{Measure:} flat $\dd r$ $\to$ throat weight $w(r)=R/r$
  from $\sqrt{g_{rr}}$ on the AdS-like confining metric of~\cite{NuclearDecay2026}.
\item \textbf{Alpha trial:} Gaussian/HO $\to$ Woods--Saxon primary
  $u_\alpha(r)=r\bigl(1+\ee^{(r-R_\alpha)/a_{\mathrm{WS}}}\bigr)^{-1}$,
  with optional WS$\times$Coulomb-tail match for comparison.
\item \textbf{Artifacts:} new code
  \texttt{palpha\_overlap\_throat.py}, JSON
  \texttt{palpha\_overlap\_throat\_results.json}; baseline proxy files
  retained.
\end{itemize}

\subsection{What remains open}
\begin{itemize}
\item Isotope-specific spectroscopic amplitudes beyond one global $S$.
\item Self-consistent bound-state solution of WS+Coulomb (R-matrix /
  cluster eigenmode) rather than a phenomenological trial.
\item Overlap constructions that place non-vanishing flag amplitude in
  the barrier $[R,b]$ (the present Dirichlet $u_{\mathrm{in}}$ vanishes
  at $R$).
\item Absolute uniqueness of preformation from many-body structure.
\end{itemize}

\begin{remark}[Next]
Parametric isotope-resolved $S(A,Z)$ (with leave-one-out) and a
self-consistent WS+Coulomb interior eigenmode are implemented as
comparison channels in the refined sequel
\texttt{flag\_condensate\_palpha\_refined.tex} /
\texttt{palpha\_overlap\_refined.py}.  On the stated set, parametric
$S(A,Z)$ reduces LOO residual RMS most among non-tautological protocols;
see that note for metrics.
\end{remark}

\section{Throat weight and overlap definition}

Companion metric~\cite{NuclearDecay2026}:
\begin{equation}
\dd s^2 = \frac{R^2}{r^2}\,\dd r^2 + \frac{r^2}{R^2}\eta_{\mu\nu}\,\dd x^\mu\dd x^\nu.
\end{equation}
We take the radial measure factor $\sqrt{g_{rr}}=R/r$ and define
\begin{equation}
P_{\mathrm{model}}
=
\frac{\bigl|\displaystyle\int_0^R u_{\mathrm{in}}(r)\,u_\alpha(r)\,w(r)\,\dd r\bigr|^2}
{\displaystyle
\Bigl(\int_0^R u_{\mathrm{in}}^2\,w\,\dd r\Bigr)
\Bigl(\int_0^R u_\alpha^2\,w\,\dd r\Bigr)},
\qquad w(r)=\frac{R}{r}.
\end{equation}
For reduced modes $u\sim O(r)$ near the origin the integrands remain
integrable.  Flat $w\equiv 1$ is recomputed for side-by-side comparison
(including reproduction of~\cite{PalphaOverlap2026}).

\begin{remark}[Domain choice]
Overlap support is $[0,R]$.  The Coulomb barrier interval $[R,b]$
determines Gamow $W$ in~\cite{NuclearDecay2026}; it is not used here as
overlap support for the Dirichlet interior mode.
\end{remark}

\section{Radial modes}

\subsection{Confined flag mode ($\ell=0$)}
Unchanged:
\begin{equation}
u_{\mathrm{in}}(r)=r\,j_0(kr),\qquad k=\pi/R.
\end{equation}

\subsection{Alpha trials}
Primary Woods--Saxon cluster packet (inputs $r_0=1.2\,\mathrm{fm}$,
$a_{\mathrm{WS}}=0.55\,\mathrm{fm}$):
\begin{equation}
u_\alpha^{\mathrm{WS}}(r)
=
\frac{r}{1+\ee^{(r-R_\alpha)/a_{\mathrm{WS}}}},
\qquad
R_\alpha=r_0\,4^{1/3}\approx 1.905\,\mathrm{fm}.
\end{equation}
Comparison: Gaussian/HO as in~\cite{PalphaOverlap2026}; optional
WS with exponential$\times R_m/r$ tail matched in value at
$R_m=R_\alpha+2a_{\mathrm{WS}}$ ($\kappa=1/a_{\mathrm{WS}}$,
phenomenological).

\section{Numerical protocol}

Simpson quadrature, $N=4096$ on $[0,R]$.  Implementation:
\texttt{rh\_papers\_may21/acs-framework/code/palpha\_overlap/palpha\_overlap\_throat.py}.
Results:
\texttt{docs/palpha\_overlap\_throat\_results.json}.
Logs:
\texttt{docs/palpha\_overlap\_throat\_logs/}.

\section{Results on the stated set}

\begin{table}[H]
\centering
\caption{\textbf{Throat+WS overlap vs extracted $\log_{10} P_\alpha$.}
Primary model; $S$ is one global factor.}
\label{tab:throat-overlap}
\vspace{0.2em}
\scriptsize
\renewcommand{\arraystretch}{0.75}
\setlength{\tabcolsep}{5pt}
\begin{tabular}{lcccccc}
\toprule
\textbf{Isotope} & \textbf{$R$ (fm)} & \textbf{$\log_{10} P_{\mathrm{model}}$}
 & \textbf{$\log_{10} P_\alpha^{\mathrm{ext}}$}
 & \textbf{resid.\ (raw)}
 & \textbf{$\log_{10}(S P_{\mathrm{model}})$}
 & \textbf{resid.\ ($S$)} \\
\midrule
$^{212}\mathrm{Po}$ & 8.99 & $-0.4441$ & $-1.53$ & $+1.0859$ & $-1.9419$ & $-0.4119$ \\
$^{214}\mathrm{Po}$ & 9.02 & $-0.4465$ & $-1.19$ & $+0.7435$ & $-1.9443$ & $-0.7543$ \\
$^{216}\mathrm{Po}$ & 9.05 & $-0.4489$ & $-1.04$ & $+0.5911$ & $-1.9467$ & $-0.9067$ \\
$^{218}\mathrm{Po}$ & 9.07 & $-0.4505$ & $-0.82$ & $+0.3695$ & $-1.9483$ & $-1.1283$ \\
$^{220}\mathrm{Rn}$ & 9.10 & $-0.4529$ & $-1.12$ & $+0.6671$ & $-1.9507$ & $-0.8307$ \\
$^{224}\mathrm{Ra}$ & 9.16 & $-0.4576$ & $-1.54$ & $+1.0824$ & $-1.9555$ & $-0.4155$ \\
$^{226}\mathrm{Ra}$ & 9.18 & $-0.4592$ & $-1.73$ & $+1.2708$ & $-1.9571$ & $-0.2271$ \\
$^{228}\mathrm{Th}$ & 9.21 & $-0.4616$ & $-1.87$ & $+1.4084$ & $-1.9595$ & $-0.0895$ \\
$^{232}\mathrm{Th}$ & 9.26 & $-0.4656$ & $-2.17$ & $+1.7044$ & $-1.9634$ & $+0.2066$ \\
$^{234}\mathrm{U}$  & 9.29 & $-0.4679$ & $-2.64$ & $+2.1721$ & $-1.9658$ & $+0.6742$ \\
$^{238}\mathrm{U}$  & 9.34 & $-0.4718$ & $-1.97$ & $+1.4982$ & $-1.9697$ & $+0.0003$ \\
$^{238}\mathrm{Pu}$ & 9.34 & $-0.4718$ & $-3.28$ & $+2.8082$ & $-1.9697$ & $+1.3103$ \\
$^{240}\mathrm{Pu}$ & 9.36 & $-0.4734$ & $-2.91$ & $+2.4366$ & $-1.9713$ & $+0.9387$ \\
$^{244}\mathrm{Cm}$ & 9.42 & $-0.4781$ & $-3.61$ & $+3.1319$ & $-1.9760$ & $+1.6340$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Primary summary (throat + Woods--Saxon)}
\begin{align*}
\langle\log_{10} P_{\mathrm{model}}\rangle &= -0.4607
\quad(\sigma=0.0110),\\
\langle\log_{10} P_\alpha^{\mathrm{ext}}\rangle &= -1.9586
\quad(\sigma=0.8656),\\
r\bigl(\log_{10} P_{\mathrm{model}},\,\log_{10} P_\alpha^{\mathrm{ext}}\bigr)
&= +0.8875,\\
\mathrm{RMS}_{S=1} &= 1.7099,\qquad
S_{\star}=3.178\times 10^{-2},\qquad
\mathrm{RMS}_{S_\star}=0.8247.
\end{align*}

\subsection{Model comparison on the stated set}
\begin{table}[H]
\centering
\caption{Flat vs throat measure; Gaussian vs WS alpha trial.}
\label{tab:model-compare}
\scriptsize
\renewcommand{\arraystretch}{0.8}
\begin{tabular}{lcccc}
\toprule
\textbf{Model} & $\langle\log_{10} P\rangle$ & $r$ & $\mathrm{RMS}_{S=1}$ & $\mathrm{RMS}_{S_\star}$ \\
\midrule
throat + WS (primary) & $-0.4607$ & $+0.8875$ & $1.7099$ & $0.8247$ \\
throat + WS$\times$Coulomb tail & $-0.4825$ & $+0.8875$ & $1.6908$ & $0.8245$ \\
throat + Gaussian & $-0.2791$ & $+0.8881$ & $1.8715$ & $0.8260$ \\
flat + WS & $-0.6786$ & $+0.8876$ & $1.5201$ & $0.8200$ \\
flat + Gaussian (baseline proxy) & $-0.3905$ & $+0.8882$ & $1.7705$ & $0.8220$ \\
\bottomrule
\end{tabular}
\end{table}

Relative to flat+Gaussian~\cite{PalphaOverlap2026}:
$\Delta\langle\log_{10} P\rangle=-0.0702$,
$\Delta r=-0.0007$,
$\Delta\mathrm{RMS}_{S_\star}=+0.0027$,
$S$ moves from $2.70\times 10^{-2}$ to $3.18\times 10^{-2}$.

\begin{remark}[Interpretation]
On this set, warped measure and WS shape shift absolute scale and
slightly compress $\sigma(\log_{10} P_{\mathrm{model}})$, but do not
materially change correlation with extracted $\log_{10} P_\alpha$.
Isotope-specific residual structure remains; one global $S$ absorbs mean
offset only.
\end{remark}

\section{Conclusion}
We have advanced the flat $[0,R]$ proxy of~\cite{PalphaOverlap2026} to
an AdS-like throat weight $w=R/r$ with Woods--Saxon alpha trial on the
same 14-isotope set.  Primary metrics:
$\langle\log_{10} P\rangle=-0.4607$, $r=+0.8875$,
$S_\star\approx 0.032$.  Correlation is essentially unchanged versus the
flat-Gaussian proxy.  Isotope-resolved / parametric $S$ and
self-consistent WS+Coulomb eigenmodes are reported in the refined
sequel~\cite{PalphaRefined2026}.  Absolute uniqueness is not claimed.

\begin{thebibliography}{99}
\bibitem{NuclearDecay2026}
  ACS Theoretical Physics Working Group,
  \emph{Nuclear Decay as a Spectral Bifurcation of the Flag Condensate},
  ACS Framework note (2026),
  \texttt{flag\_condensate\_nuclear\_decay.tex} /
  \texttt{papers/notes/Flag\_Condensate\_Nuclear\_Decay.tex}.
\bibitem{PalphaOverlap2026}
  ACS Theoretical Physics Working Group,
  \emph{Standing-Wave Overlap Proxy for Alpha Preformation},
  ACS Framework note (2026),
  \texttt{flag\_condensate\_palpha\_overlap.tex} /
  \texttt{papers/notes/Flag\_Condensate\_Palpha\_Overlap.tex}.
\bibitem{PalphaRefined2026}
  ACS Theoretical Physics Working Group,
  \emph{Refined $P_\alpha$ Overlap: Parametric $S$ and WS+Coulomb Eigenmode},
  ACS Framework note (2026),
  \texttt{flag\_condensate\_palpha\_refined.tex} /
  \texttt{papers/notes/Flag\_Condensate\_Palpha\_Refined.tex}.
\bibitem{gamow1928}
  G. Gamow, \emph{Zur Quantentheorie des Atomkernes},
  Z. Phys. \textbf{51}, 204 (1928).
\end{thebibliography}

\end{document}
