% Co-governed and enforced under the Sovereign Integrity Protocol License (SIP License v1.1):
% https://github.com/tensorrent/ACS-Framework/blob/main/LICENSE
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,urlcolor=blue!50!black,citecolor=blue!50!black}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage[protrusion=true,expansion=false]{microtype}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  framesep=4pt,
  backgroundcolor=\color{gray!6},
  columns=fullflexible,
  keepspaces=true,
}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Form/Function Relativity}
\fancyhead[R]{\small Form/Function Relativity}
\fancyfoot[C]{\thepage}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{plain}
\newtheorem{principle}{Principle}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newcommand{\real}{\mathrm{real}}

\title{\textbf{Form/Function Relativity:\\
The Reference Frame Is the Perspective}}
\author{B.\ Wallace \quad (TensorRent)\\ \small with computational verification by Claude (Anthropic)}
\date{\small Companion to \textit{Spectral Rigidity and Shuffled Spacing Discriminant}\ (FF06e)\ ---\ SIP License v1.1}

\begin{document}
\maketitle

\begin{abstract}
The shuffle knife (FF06e) sorts each spectral witness into \emph{form} (its value is a property of the distribution class) or \emph{function} (its value reads the object, $\zeta$, specifically) by comparing the real zeros against a marginal-matched surrogate. That verdict is stated as if it were a property of the witness. It is not. It is a property of the \emph{pair} (witness, reference measure), and the reference measure --- the surrogate ensemble against which the witness is scored --- is a free choice. Change the choice and the verdict moves. We make this operational by scoring the same five witnesses against a \emph{ladder} of reference frames ordered by how much structure each preserves: Poisson (no level repulsion), the GUE-marginal shuffle (nearest-neighbour spacing law only), a full-GUE bulk ensemble (the two-point correlation), and $\zeta$ itself (the arithmetic). Refining the frame turns witnesses from function to form one structural class at a time --- a staircase. The repulsion and Wigner-shape witnesses are \textbf{function} against Poisson ($44$--$50\sigma$) and \textbf{form} against the GUE-marginal frame ($0.1$--$0.2\sigma$): the identical witness carries opposite labels in the two frames. The rigidity (number-variance) witness is function against both of those and form ($1.6\sigma$) only once the frame preserves the two-point function. The arithmetic prime-resonance witness is function against every frame short of $\zeta$ itself ($260$--$832\sigma$): it is the irreducibly object-specific content. The form/function label is therefore a coordinate on (witness $\times$ frame), not an intrinsic attribute --- \emph{one witness's form is another witness's function}. This is the spectral instance of a general role-relativity in nested systems, the same statement carried at different depths by \emph{one man's trash is another's treasure} and \emph{one party's security is another's leverage}: a component's role is conferred by the enclosing system that resolves it, and a differently-resolving system in the nest confers the opposite role. It is also the same reference measure $\mathrm{d}\mu/\mathrm{d}\nu$ that already carries the perspective in the BCH--transfer-entropy lemma of the core monograph; here it is exhibited numerically. Reproducible at seed $20260423$. Nothing here proves RH; negatives are kept first-class.
\end{abstract}

\section{The observation}\label{sec:obs}

FF06e defines form and function through a single surrogate ensemble.

\begin{definition}[Form and function, FF06e Def.~2, recalled]\label{def:ff-single}
A witness $W$ is a scalar functional of a spectrum. Evaluate it on the real zeros, $W_\real$, and on an ensemble of marginal-matched surrogates (the unfolded spacings permuted and re-accumulated), with mean $\overline{W}$ and standard deviation $\sigma_W$. Its \emph{influence} is $z(W)=|W_\real-\overline{W}|/\sigma_W$; $W$ is \textbf{form} if $z(W)$ is small ($\lesssim 3$) and \textbf{function} if $z(W)$ is large.
\end{definition}

The definition names one surrogate --- the gap-shuffle --- and reads off a label. But the gap-shuffle is a choice, and FF06e's own scope section already records that the choice is load-bearing: the lag-1 verdict swings by $\sim\!6\sigma$ ``induced purely by the choice of GUE pipeline,'' with two legitimate GUE references disagreeing by more than the candidate signal. If replacing one repulsion-matched reference by another moves the verdict, then replacing a repulsion-matched reference by a \emph{non}-repulsion-matched one (Poisson) will move it much further. The label was never attached to the witness alone. It was attached to the witness \emph{as seen from} a reference frame, and FF06e fixed the frame silently.

We make the frame a variable. Read the other way, Definition~\ref{def:ff-single} is already the folk observation that one man's trash is another's treasure, restricted to spectra and made quantitative: the $z$-score is the amount by which one enclosing system (the surrogate) fails to resolve what another (the object) contains, and ``form'' versus ``function'' is which side of that gap the witness is being read from.

\section{Relative form and function}\label{sec:rel}

\begin{definition}[Form/function relative to a reference frame]\label{def:ff-rel}
Let $\nu$ be a reference ensemble (a ``frame'') on spectra: a distribution from which surrogates are drawn. For a witness $W$ define
\[
  z_\nu(W)\;=\;\frac{\big|W_\real-\mathbb{E}_\nu[W]\big|}{\sqrt{\operatorname{Var}_\nu[W]}}.
\]
$W$ is \textbf{function relative to $\nu$} if $z_\nu(W)>\tau$ (we use $\tau=3$) and \textbf{form relative to $\nu$} otherwise. Definition~\ref{def:ff-single} is the special case $\nu=$ gap-shuffle.
\end{definition}

The verdict is now explicitly a two-argument object. The content of this note is that the second argument matters, and matters in a structured way. Before specialising to spectra, we state the shape of the claim at the level at which it is really operating.

\begin{principle}[Role-relativity in nested systems]\label{prin:nested}
Let a component sit inside a family of systems ordered by \emph{nesting}: system $\nu_1$ is nested in $\nu_2$ (written $\nu_1\preceq\nu_2$) when every structure $\nu_1$ resolves is also resolved by $\nu_2$. Each enclosing system reads off a two-valued \emph{role} for the component --- the label it assigns given what that system can and cannot resolve. Then the role is a coordinate on (component $\times$ system), not a property of the component: the same component is assigned opposite roles by different systems in the nest, and ascending the nest can only convert the \emph{marked} role (the one that records unresolved structure) into the \emph{unmarked} one, never the reverse. Each component therefore has a single threshold level at which its role flips, once, and stays flipped. \emph{One man's trash is another's treasure; one party's security is another's leverage; one field's form is another's function} --- three readings of a single component by three enclosing systems that resolve it differently.
\end{principle}

The spectral case is the instance of this principle in which the components are witnesses, the role is $\{$form, function$\}$, and the nest is the refinement ladder of \S\ref{sec:ladder}.

\begin{principle}[Form/function relativity]\label{prin:rel}
The label $\{$form, function$\}$ is a coordinate on the product space (witness $\times$ frame), not a property of the witness. For a fixed witness $W$ there exist frames $\nu_1,\nu_2$ with $W$ \emph{form} relative to $\nu_1$ and \emph{function} relative to $\nu_2$; equivalently, relative to a fixed frame two witnesses split into form and function, and the split is not preserved when the frame changes. \emph{One witness's form is another witness's function.}
\end{principle}

The mechanism is not mysterious. A witness is function relative to $\nu$ exactly when it reads structure that $\nu$ does \emph{not} reproduce. So the label tracks the difference between what the witness reads and what the frame preserves --- a relation between two things, neither of which is the witness by itself. Order the frames by refinement (how much structure each preserves) and the labels must fall monotonically:

\begin{proposition}[Monotone staircase]\label{prop:stair}
Let $\nu_1 \preceq \nu_2$ mean that every scalar structure preserved (in ensemble mean) by $\nu_1$ is also preserved by $\nu_2$. If a witness $W$ reads only structure that $\nu_2$ preserves, then $z_{\nu_2}(W)\le z_{\nu_1}(W)$ up to sampling noise: refining the frame can turn function into form but never the reverse. Hence along a refinement chain each witness has a single threshold frame at which it falls from function to form, and the witnesses are totally ordered by the depth of structure they read.
\end{proposition}

This is a statement about the geometry of the (witness $\times$ frame) table, and it is testable: the table should be a staircase, not a scatter. It is also Principle~\ref{prin:nested} made a theorem: the staircase is ``trash becomes treasure'' with the enclosing systems named --- a component keeps its marked role until the system around it is large enough to resolve what it reads, then flips exactly once. The relabelling is not a paradox to be resolved but a coordinate to be read off.

\section{The frame ladder}\label{sec:ladder}

We instantiate a four-rung refinement chain on the Riemann zeros. Each rung preserves strictly more structure than the last.

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}llp{6.6cm}@{}}
\toprule
Frame $\nu$ & preserves & construction \\
\midrule
Poisson            & density only        & i.i.d.\ exponential spacings, density-matched \\
GUE-marginal       & $+$ NN-spacing law  & permute the real unfolded spacings, re-accumulate \\
GUE-full (bulk)    & $+$ two-point corr. & eigenvalues of a random Hermitian matrix, bulk unfolded \\
$\zeta$ itself     & $+$ arithmetic      & the real zeros (terminal frame) \\
\bottomrule
\end{tabular}
\end{center}

\noindent The chain is Poisson $\preceq$ GUE-marginal $\preceq$ GUE-full $\preceq \zeta$. This is a \emph{nest} of systems in the sense of Principle~\ref{prin:nested}: each frame is an enclosing system that resolves strictly more of the spectrum's structure than the one before, so each reads a witness's form/function role differently, and the everyday dualities (trash/treasure, security/leverage) are the same relabelling one nesting level apart. Poisson destroys level repulsion; the GUE-marginal shuffle restores the exact nearest-neighbour spacing distribution but destroys all correlations of order $\ge2$ and all arithmetic; the full-GUE bulk restores the RMT two-point function but still carries no arithmetic; $\zeta$ itself is the only frame that carries the primes, and against it every witness is trivially form ($z=0$ against itself). The reference measure $\nu$ is exactly the free reference $\mathrm{d}\mu/\mathrm{d}\nu$ of the BCH--transfer-entropy lemma (core monograph, Lemma~2.9): the information asymmetry, and with it the form/function reading, is always measured against a \emph{chosen} $\nu$. Choosing $\nu$ differently relabels; that choice is what we call the perspective.

We score five witnesses, ordered by the depth of structure they read: \emph{repulsion} (fraction of spacings below half the mean) and \emph{shape} ($L^2$ distance of the spacing histogram to the GUE Wigner surmise), both reading only the nearest-neighbour spacing law; \emph{lag-1} serial correlation, reading local order; \emph{rigidity} (number variance in fixed windows), reading the two-point correlation; and \emph{arith} ($\sum_{p\le 29}\langle\cos(\gamma\log p)\rangle^2$), reading the arithmetic.

\section{Measurement}\label{sec:meas}

On the first $N=4000$ Odlyzko zeros at seed $20260423$, with $200$ surrogate draws for the two exact frames and $40$ for GUE-full (bulk, central $80\%$ trimmed before unfolding). Cells give $z_\nu$ and the resulting label at $\tau=3$.

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lcccl@{}}
\toprule
witness & vs Poisson & vs GUE-marginal & vs GUE-full & reads \\
\midrule
repulsion & $49.5$ \textsc{func} & $\mathbf{0.2}$ \textsc{\textbf{form}} & $5.3$ \textsc{func} & NN-spacing law \\
shape     & $44.3$ \textsc{func} & $\mathbf{0.1}$ \textsc{\textbf{form}} & $6.5$ \textsc{func} & NN-spacing law \\
lag-1     & $25.7$ \textsc{func} & $25.9$ \textsc{func} & $6.5$ \textsc{func} & local order \\
rigidity  & $10.7$ \textsc{func} & $10.4$ \textsc{func} & $\mathbf{1.6}$ \textsc{\textbf{form}} & two-point corr. \\
arith     & $259.9$ \textsc{func} & $832.2$ \textsc{func} & $479.1$ \textsc{func} & arithmetic \\
\bottomrule
\end{tabular}
\end{center}

\noindent Read the rows. \textbf{Repulsion and shape} are the theorem in miniature: $44$--$50\sigma$ function against Poisson, $0.1$--$0.2\sigma$ form against the GUE-marginal frame. The identical scalar functional, on the identical zeros, is function in one frame and form in the next --- Principle~\ref{prin:rel} made numerical. Their agreement with the critical line, viewed from Poisson, looks like a sighting of $\zeta$; viewed from a repulsion-matched frame it is revealed as a property of the class that any GUE-spaced sequence would reproduce. \textbf{Rigidity} falls one rung later: function against Poisson and against the spacing-marginal shuffle (both fail to reproduce the two-point function), form ($1.6\sigma$) only once the frame is full GUE. \textbf{Arith} never falls: it is function against every frame that lacks the primes, and its $z$ is two to three orders of magnitude above threshold. It is the object-specific content that no non-arithmetic frame can absorb --- the explicit-formula coupling, i.e.\ unique factorisation read spectrally.

The table is a staircase, as Proposition~\ref{prop:stair} requires: reading down the ``threshold frame'' at which each witness turns form gives repulsion/shape at GUE-marginal, rigidity at GUE-full, arith beyond every frame short of $\zeta$. The witnesses are thereby \emph{ordered} by the depth of structure they read, and that order is an output of the measurement, not an input.

The two exact frames (Poisson, GUE-marginal) run in $\sim\!2$\,s and carry the load-bearing flip; the GUE-full column requires dense diagonalisation and is gated behind \texttt{-{}-full} ($\sim\!7$\,min). The driver is \texttt{code/hp\_knife\_suite/hp\_form\_function\_relativity.py}.

\section{Why this is the same object as the monograph's asymmetry}\label{sec:mono}

The relativity is not a quirk of the knife; it is the discrete face of a structure already central to the framework.

\textbf{The reference measure is the perspective.} The BCH--transfer-entropy lemma writes the information asymmetry as $\Delta\mathcal{I}(\varepsilon)=\varepsilon\,\langle f-g,\nabla\log\tfrac{\mathrm{d}\mu}{\mathrm{d}\nu}\rangle_\mu + 2\varepsilon^2\langle[f,g],\nabla\log\tfrac{\mathrm{d}\mu}{\mathrm{d}\nu}\rangle_\mu+\cdots$, against a reference measure $\nu$. The Radon--Nikod\'ym derivative $\mathrm{d}\mu/\mathrm{d}\nu$ is exactly the frame of Definition~\ref{def:ff-rel}: what counts as asymmetry, and therefore what counts as function, is measured against a chosen $\nu$. The knife's ``surrogate ensemble'' and the lemma's ``reference measure'' are the same slot in the same formula. Form/function relativity is what that slot being free \emph{looks like} on a spectrum.

\textbf{Form and Function were already frame-relative in gauge theory.} The monograph assigns Form to the observable curvature and Function to the potential (the vierbein/connection pair, the field-strength/gauge-field pair). That assignment is itself gauge-dependent: which half of $e^{i\theta}=\cos\theta+i\sin\theta$ is ``the observable'' depends on the phase origin, and a gauge rotation trades one for the other. The core monograph's own remark that ``the distinction is not merely labelling'' is correct \emph{within} a fixed frame and incomplete across frames: within a frame Form constrains Function and vice versa; across frames the assignment rotates. Principle~\ref{prin:rel} is the number-theoretic instance of that rotation --- the same statement the title of the parent programme already makes, that form and function are generated by the asymmetry of a pair and not carried by either member alone.

\section{What this establishes, and where it stops}\label{sec:scope}

\textbf{Establishes.} The form/function verdict of FF06e is frame-relative in the precise sense of Definition~\ref{def:ff-rel}: on the first $4000$ zeros the repulsion and shape witnesses are function against Poisson and form against a repulsion-matched frame, and rigidity is function until the frame preserves the two-point function. The (witness $\times$ frame) table is a monotone staircase, so the witnesses are totally ordered by the depth of structure they read, with the arithmetic witness strictly deepest. This sharpens FF06e's single-surrogate categorisation into a frame-indexed one and identifies the frame with the reference measure of the monograph's asymmetry lemma.

\textbf{Does not establish.} Nothing here proves the Riemann Hypothesis, constructs a Hilbert--P\'olya operator, or bounds off-line behaviour. ``Function relative to $\nu$'' means ``reads structure absent from $\nu$,'' not ``proves a property of $\zeta$.'' That the arithmetic witness stays function against every non-arithmetic frame is consistent with, but not evidence for, RH: it says only that the primes are not reproduced by any GUE-type null, which is the explicit formula, not a theorem about zero locations. The general role-relativity (Principle~\ref{prin:nested}) and the everyday dualities are framing, not measurement: the one instance measured here is the Riemann spectrum (\S\ref{sec:meas}, tier T1), and the folk sayings are cited to locate the principle, not as evidence for it.

\textbf{Two honest edges.} (i) The GUE-full frame is a finite-$N$, unfolded ensemble; its column is approximate, and it is the two exact frames that carry the argument. Consistent with that approximation, lag-1 remains nominally function ($6.5\sigma$) against GUE-full where Proposition~\ref{prop:stair} would place its threshold --- a finite-$N$ apparatus effect of the same kind FF06e flagged (the $\sim\!6\sigma$ inter-pipeline swing), not a claimed beyond-GUE signal. (ii) The threshold $\tau=3$ is a convention; the staircase structure, not the exact cell labels near the threshold, is the result. Both edges are recorded so the table is not over-read.

\vspace{1em}
\noindent\rule{\linewidth}{0.4pt}\\
\small This companion accompanies \textit{Spectral Rigidity and Shuffled Spacing Discriminant} (FF06e) and \textit{Spectral Witness Survival}. All measurements use the real first Riemann zeros and the canonical seed $20260423$; the relativity suite is reproducible standalone via \texttt{hp\_form\_function\_relativity.py}. SIP License v1.1.

\end{document}
