\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,urlcolor=blue!50!black,citecolor=blue!50!black}
\usepackage{fancyhdr}
\usepackage[protrusion=true,expansion=false]{microtype}
\pagestyle{fancy}\fancyhf{}
\fancyhead[L]{\small The Reversible Flattening}\fancyhead[R]{\small The Reversible Flattening}
\fancyfoot[C]{\thepage}
\newtheorem{theorem}{Theorem}\newtheorem{proposition}{Proposition}
\newtheorem{criterion}{Criterion}\newtheorem{definition}{Definition}
\newtheorem{observation}{Observation}

\title{\textbf{The Reversible Flattening:\\
Algebraic Representation as a Recoverable Compression of Relational Structure}\\[4pt]
\large A homomorphism criterion for when scalars should be returned to geometry}
\author{B.\ Wallace \quad (TensorRent)\\
\small with computational verification by Claude (Anthropic)}
\date{\small Parent of \textit{The Geometry Engine} / \textit{When a Number Lies} \ ---\ SIP License v1.1 \ ---\ Seed \texttt{20260423}}

\begin{document}
\maketitle

\begin{abstract}
We advance a single mathematical thesis with a precise, testable criterion, and characterize the boundary it
runs into. Much of arithmetic and algebra is a \emph{flattening}: a relational object (a factorization, a
divisor lattice, a dominance relation) is projected onto a scalar or a symbol string because, historically,
symbols on paper were the computable medium. We give the exact condition under which that flattening is
\emph{reversible} --- recoverable losslessly to a relational form one can compute in directly --- and show it
is the condition that the operation producing the scalar be a \emph{homomorphism} from the object's structure
into a structure the relational representation stores. The boundary of reversibility (the ``seam'') is not a
void: we characterize it precisely as \emph{convolution} --- the operation that adds indices while
multiplying values --- and verify that the additive structure it was thought to lack is real but
\emph{cardinal} (the partition function, the size of the cloud) rather than \emph{unique} (a factorization,
a point), and is encoded multiplicatively by its generating function (Euler's product, verified). Two kinds
of structure --- unique and cardinal --- meet in convolution, which is crossable for density and uncrossable
for identity; this single fact organizes the explicit formula, the difficulty of fully homomorphic
encryption, and the limits below. Under this criterion, three regimes appear and are each verified. (1) The
\emph{homomorphic} regime --- multiplication, $\gcd$, $\mathrm{lcm}$, powers, and every multiplicative
arithmetic function --- is reversible: by the Fundamental Theorem of Arithmetic, $(\mathbb{Z}_{>0},\times)$
is isomorphic to $(\bigoplus \mathbb{N},+)$ on exponent vectors, and we verify exact computation across seven
applied-mathematics domains, $14/14$ tests exact against ground truth, including objects of $\sim\!5000$
digits that overflow the scalar form. (2) The \emph{non-transitive} regime --- dominance relations, social
choice --- admits \emph{no} faithful scalar at all: a Condorcet cycle forces every total order to contradict
the majority relation, so the relation is mandatory, not merely preferable, and the cyclicity persists at
scale. (3) The \emph{seam} --- addition and ordering --- is non-homomorphic on factor structure and therefore
irreversible: it is the genuine, fundamental wall, proven by direct counterexample. The thesis is thus not
``geometry is better than algebra'' but the sharper claim that the algebraic flattening was a medium-driven
compression, reversible exactly on the homomorphic and relational classes, and that modern computation lets
us both perform and \emph{verify} the inverse transposition. We keep the full negative ledger --- two
falsified search heuristics and four architecture-integration hypotheses falsified against live source ---
because the boundary, not the enthusiasm, is the result.
\end{abstract}

\section{Thesis}

Mathematics began with observation and construction --- lengths, angles, areas --- and was then abstracted
into symbol and number because that form was mechanically manipulable by hand. This abstraction was an
enormous gain and also a \emph{compression}: it flattened relational content into coordinate- and
digit-strings, and a great deal of later effort (managing precision, overflow, coordinate artifacts) is the
cost of recovering relations the flattening discarded. The thesis of this paper is precise:

\begin{quote}
The algebraic flattening of a relational object into a scalar is \emph{losslessly reversible} if and only if
the operation that produces the scalar is a homomorphism from the object's structure into a structure the
relational representation stores directly. Where reversible, computing in the restored relational form is
more faithful and frequently more tractable; where not, the scalar is native and the flattening is genuine.
\end{quote}

This is a statement about \emph{representation}, not about the superiority of one notation. It explains a
historical contingency (symbols were the computable medium) and supplies a criterion that did not previously
have a name, then verifies the criterion across applied mathematics.

\section{The criterion}

\begin{definition}[Flattening and its inverse]
A \emph{flattening} of a structured object $R$ is a map $\pi:R\to s$ to a scalar or symbol string. It is
\emph{reversible} if there is a representation $\rho(R)$ --- the relational form --- and a recovery
$\rho(R)\to s$ that is exact, such that the operations of interest are computed on $\rho(R)$ without forming
$s$.
\end{definition}

\begin{criterion}[Homomorphism]
The flattening of an operation $\star$ is reversible exactly when $\star$ is a homomorphism
$(\text{objects},\star)\to(\text{stored relational structure},\oplus)$. Then $\rho(a\star b)=\rho(a)\oplus
\rho(b)$ and the relational form computes $\star$ without forming the scalar.
\end{criterion}

For integers under multiplication this is the Fundamental Theorem of Arithmetic: factorization is an
isomorphism $(\mathbb{Z}_{>0},\times)\cong(\bigoplus_p\mathbb{N},+)$, so multiplication becomes vector
addition of exponents, $\gcd$ and $\mathrm{lcm}$ become coordinatewise $\min$ and $\max$, powers become
scaling. We verify $\rho(ab)=\rho(a)\oplus\rho(b)$ directly. Addition fails the criterion: $\rho(a+b)$ is not
a function of $\rho(a),\rho(b)$ in any structure-preserving way --- $360+84=444$ and $363+81=444$ have
identical sums from disjoint summand structures, and a unit change of input ($360+86=446$) yields an
unrelated factor structure. Addition is not a homomorphism on factor structure; its flattening is
irreversible. This is the seam, now stated as the exact non-homomorphic boundary rather than as a list of
hard cases.

\section{Regime 1: the homomorphic class, verified exact}

Every operation that is homomorphic into exponent-vector structure is reversible, and we computed each in the
relational form, exact against ground truth (oracle: standard libraries; seed \texttt{20260423}).

\begin{center}
\begin{tabular}{lll}
\toprule
domain & operations & verification \\
\midrule
arithmetic functions & $d(n),\sigma(n),\sigma_2(n),\varphi(n),\mu(n)$ & exact, $300$ values each \\
$\gcd$ / $\mathrm{lcm}$ & $\min$/$\max$ on exponents & exact, $300$ pairs each \\
binomial / factorial & Legendre exponent formula & $\binom{20}{7},\binom{100}{50},\binom{500}{123}$ exact \\
modular products & homomorphism preserved under $\bmod$ & concentrated chain $\bmod\,(2^{61}{-}1)$ exact \\
lattice index / SNF & product of elementary divisors & $\sim\!1479$-digit index, exact, no overflow \\
multinomial multiplicity & $N!/\prod n_i!$ & $\sim\!4978$-digit $W$, exact \\
radical / perfect power & squarefree kernel, exponent $\gcd$ & exact, $200$ values \\
\bottomrule
\end{tabular}
\end{center}

\begin{observation}[$14/14$ exact; the flattening visibly fails where the relation does not]
All fourteen tests are exact. Two cases are decisive for the thesis: the multinomial $W$ ($\sim\!4978$
digits) and the product of the first $5000$ primes ($\sim\!20975$ digits) are computed exactly in relational
form, while \emph{forming the scalar} exceeded the host language's integer-to-string limit --- the flattening
hit a wall the relational representation did not. The headline tractability instance (companion FF06g/h) is
an exact-optimal search for the integer maximizing $d(n)$ that runs in exponent space to $N=10^{80}$,
verified-optimal against brute force at four anchors, sixty-two orders of magnitude past enumeration.
\end{observation}

\section{Regime 2: the non-transitive class, where no scalar exists}

The strongest form of the thesis is not lossiness but impossibility.

\begin{theorem}[No faithful scalar under cycles]
If the pairwise majority relation contains a cycle $A\succ B\succ C\succ A$, every total order contradicts at
least one majority verdict. No scalar ranking represents the relation.
\end{theorem}

\begin{proof}[Verification]
Exhaustive over all orderings of the Condorcet instance: the minimum number of contradicted majority edges
is one, not zero. Total orders are transitive; the majority relation is not; a transitive structure cannot
realize a cycle. This is the social-choice form of the same obstruction by which box containment (a partial
order) cannot encode rock-paper-scissors.
\end{proof}

\begin{observation}[Structural, not small-sample]
Measured cycle frequency rises with options ($\sim\!7\%,43\%,78\%,99\%$ for $3,5,7,10$ candidates) and
persists with electorate size ($\sim\!48\%$ at $1001$ voters, five candidates). The impossibility is a
structural feature of aggregation. Here the relational representation (a directed dominance graph,
integer and exact) is not an improvement on a scalar --- it is the only faithful object.
\end{observation}

\section{Regime 3: the seam, the irreversible wall}

Addition and ordering are non-homomorphic on factor structure (\S2) and so their flattening is irreversible:
the sum genuinely destroys multiplicative structure. This is not a limitation of the representation; it is a
fact about the operation, and it bounds the thesis from below. Every mixed object inherits it: a determinant
is a \emph{sum} of products (irreversible) even though each product is reversible; a characteristic
polynomial's coefficients are elementary symmetric \emph{sums} (irreversible) even when the eigenvalue
\emph{product} (the determinant's magnitude) is reversible. The criterion sorts these correctly without
appeal to intuition: test whether the operation is a homomorphism into the stored structure.

\section{The seam characterized: convolution (keystone)}

The seam is not a void to be avoided; it has a precise structure, and naming it sharpens the entire thesis.

\begin{theorem}[Two kinds of structure]
The multiplicative side carries \emph{unique} structure: by the Fundamental Theorem of Arithmetic a number's
factorization is a single canonical object (a point). The additive side carries \emph{cardinal} structure:
a number's additive decompositions are non-unique (a cloud), but the \emph{size} of that cloud is exact and
deep --- the partition function $p(n)$ (e.g.\ $p(100)=190{,}569{,}292$). Addition is not structureless; its
structure is the count, not an address.
\end{theorem}

\begin{theorem}[The seam is convolution, verified]
The additive (cardinal) structure is encoded multiplicatively: Euler's identity
$\sum_n p(n)x^n=\prod_{k\ge 1}(1-x^k)^{-1}$ turns the entire additive cloud into one product (verified: the
product's coefficients equal $p(n)$ for $n\le 30$). The operation joining the two sides is
\emph{convolution}: multiplying generating functions convolves coefficients,
$(a\,x^i)(b\,x^j)=ab\,x^{i+j}$, which \emph{adds} the index and \emph{multiplies} the value. Convolution is
the native inhabitant of the seam --- simultaneously additive in the index and multiplicative in the value
(verified on coefficient sequences). The seam is therefore not impassable; it is crossed by generating
functions, and the crossing operation is convolution.
\end{theorem}

\begin{observation}[Density, not identity]
Convolution crosses the seam for \emph{counting} and not for \emph{identifying}: it yields how many, never
which one. This single property organizes three facts at once. The Riemann explicit formula (companion work)
is a convolution of primes against zeros, which is why the prime/zero relationship is statistical (a
$+122\sigma$ density signal) and never an address. Fully homomorphic encryption is hard for the same reason
the seam is real: addition and multiplication share no common homomorphic structure, so a single scheme must
host both natively (lattice/LWE) and pay in noise --- there is no cheap bridge, exactly as the seam predicts.
And the additive wall of \S2 is precisely the statement that convolution gives density without identity: the
sum's factor structure is the convolution of the summands' structures, which determines a distribution, not
a point.
\end{observation}

\section{Two independent axes: order-dependence and structure-retention}

A test of the conjecture ``commutativity measures projected-away structure'' falsified the strong form and
yielded a sharper map. Non-commutativity measures retained \emph{order-dependent} structure: scalar
$+,\times$ are commutative and retain none (an index of $0$); the cross product is anti-commutative,
retaining exactly orientation (one bit, a structured sign flip); matrix products are fully non-commutative,
retaining the entire arrangement. But commutativity does \emph{not} imply absence of structure: $\gcd$ and
$\mathrm{lcm}$ are commutative and retain the full common multiplicative structure (order-\emph{independent}
structure). Order-dependence and structure-retention are therefore \emph{independent axes}. The shape engine
lives in the commutative-and-structured cell --- which is exactly why its canonical fingerprint is
path-independent (commutativity) yet information-complete (structure). The corrected law: non-commutativity
measures retained order-dependent structure only; order-independent structure is orthogonal to it.

\section{Consequence: inherited representation as recoverable debt}

The criterion gives ``inherited technical debt'' a precise mathematical reading. A representational layer
built atop the algebraic flattening inherits the flattening's irreversibility wherever the underlying
operation was non-homomorphic, and inherits a \emph{recoverable} compression wherever it was homomorphic. The
disciplined move, when seeking improvement, is therefore not only to build forward but to test whether a
lower layer was homomorphically richer than the layer built on it --- and, crucially, modern computation lets
us \emph{test} this per operation rather than assert it. The transposition back to relational form is a
hypothesis checked by the homomorphism criterion and anchored against ground truth, not a faith. We make no
claim beyond mathematics here: the criterion is the content.

\section{Negative ledger (retained)}

Falsifications are first-class. Two search heuristics were falsified en route to the verified-optimal
$d(n)$ search (a greedy method, suboptimal; a non-admissible bound that pruned true optima). Four
architecture-integration hypotheses were falsified against live source: the multiplicative-fit of an
eigenvalue ``charge'' (the real charge composes additively, by the kernel's own specification); a
geometric dual-engine router (the live constraints are ordinal thresholds and the live relation is
transitive --- no cycles); and trust and hash layers (additive and avalanche-destroyed, respectively). In
each, a clean structural resemblance was the warning sign, and the source or the anchor was the arbiter.
These are the criterion working: the homomorphism test predicts the misses as well as the hits.

\section{Scope}

The mathematics is classical (FTA, Legendre, the multiplicativity of $\sigma,\varphi,\mu$, Smith normal
form, Arrow/Condorcet). The contribution is the \emph{criterion} --- reversibility of a flattening equals
homomorphism into the stored relational structure --- and its verification across seven applied domains, the
non-transitive impossibility class, and the additive wall. The thesis does not touch primality or
prime-counting (building a prime's relational form is the factoring wall), does not bear on the Riemann
Hypothesis (a categorical gap between finite multiplicative bookkeeping and the analytic continuation of an
infinite object), and is not an arithmetic accelerator (a compiled big integer outpaces the relational form
on raw throughput; the relational advantage is exactness, overflow avoidance, and faithful representation of
the otherwise-unrepresentable). Within those bounds it is exact, reproducible, and bounded from both sides.

\vspace{1em}\noindent\rule{\linewidth}{0.4pt}\\
\small Parent of \textit{The Geometry Engine} and \textit{When a Number Lies} (the relational principle and limits). Reference
engine, dominance engine, and the reproducible ledger \texttt{verify\_ledger.py} ($14/14$ exact) accompany
this paper. SIP License v1.1. Seed \texttt{20260423}.

\end{document}
