\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{listings}
\lstset{literate={·}{{\textperiodcentered}}1 {≥}{{$\geq$}}1 {→}{{$\rightarrow$}}1 {²}{{$^2$}}1 {^}{{\textasciicircum}}1}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,urlcolor=blue!50!black,citecolor=blue!50!black}
\usepackage{fancyhdr}
\usepackage[protrusion=true,expansion=false]{microtype}
\pagestyle{fancy}\fancyhf{}
\fancyhead[L]{\small The Reversible Flattening --- Monograph}\fancyhead[R]{\small The Reversible Flattening Monograph}
\fancyfoot[C]{\thepage}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{criterion}{Criterion}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{observation}{Observation}[section]
\newtheorem{principle}{Principle}[section]

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  rulecolor=\color{gray!40},
  backgroundcolor=\color{gray!4},
  keywordstyle=\color{blue!60!black},
  commentstyle=\color{green!40!black},
  stringstyle=\color{red!50!black},
  language=Python,
  showstringspaces=false,
  columns=fullflexible,
}

\title{\textbf{The Reversible Flattening}\\[6pt]
\large A Monograph on Relational versus Scalar Representation:\\
the Homomorphism Criterion, the Convolution Seam, and a Verified Computational Record}
\author{Bradley Wallace \quad (TensorRent)\\
\small with computational verification by Claude (Anthropic)}
\date{\small Consolidates \textit{The Geometry Engine} / \textit{When a Number Lies} / \textit{The Reversible Flattening} \ ---\ SIP License v1.1 \ ---\ Seed \texttt{20260423}}

\begin{document}
\maketitle
\begin{abstract}
\noindent This monograph consolidates a single line of inquiry: that a large class of computations abstract a
relational object to a scalar, and that doing so is a reversible structural re-representation when the operation
producing the scalar is a homomorphism into a structure the relational form stores. We develop the thesis,
state and test the homomorphism criterion --- sharpened by a $30$-domain predictive test to require an
\emph{injective} homomorphism into a \emph{pre-specified} structure (homomorphism carries density; injectivity carries identity), a one-way implication whose converse is falsified by the Cantor pairing --- and characterize the boundary of reversibility (the
``seam''): convolution crosses it for aggregate density (Euler's product) but it remains uncrossable for
individual identity. We present the complete verified computational record: two reference engines in
full source, fourteen exact applied-mathematics verifications, a structure search that reaches
$10^{80}$ for divisor maximization without forming $n$ (not a general method), the non-transitive
impossibility class, and a full ledger of negative results. The relational form is more faithful (it
preserves structure the scalar discards) and exact where the scalar overflows; it is \emph{not} generally
faster than compiled big-integer arithmetic, and \emph{not} a memory compression --- for square-free or
many-distinct-prime integers it is the same size as the scalar or slightly larger (measured: primorial
$1.25\times$, prime $1.01\times$), compressing only when exponents are large. It is a structural
re-representation, not a smaller encoding. The work is
presented as a falsifiable research program: every load-bearing claim is anchored against ground truth, the
language is matched to the evidence by a four-tier scheme, and the failures are documented as carefully as
the successes because, in a study of limits, the failures are the result.
\end{abstract}

\tableofcontents
\newpage

\section{Introduction and methodology}

\subsection{The thesis in one paragraph}
Mathematics began with observation and geometric construction and was then abstracted into symbol and number
because that form was mechanically manipulable by hand. This abstraction was a great gain and also a
\emph{structural flattening} (not, in general, a size reduction --- for square-free integers the factor
representation is the same size or slightly larger than the scalar; it is smaller only under repeated
factors): it recast relational content as scalar- and digit-strings. The thesis of this work
is that the flattening is \emph{losslessly reversible} when the operation producing the scalar is a
homomorphism from the object's structure into a structure the relational representation stores directly; that
where reversible, computing in the restored relational form is more faithful and exact where the scalar
overflows (though not generally faster than compiled big-integer arithmetic); and
that the boundary of reversibility is not a void but a precise object --- convolution --- which is crossable
for density and uncrossable for identity.

\subsection{Method: adversarial compression}
Every claim followed one cycle: conjecture $\to$ pre-registered test against ground truth $\to$ a theorem
(if it survived), a sharpened negative result (if it failed), or a scoped observation (if partial). Results
carry a four-tier label: \textbf{T1} machine-verified against ground truth; \textbf{T2} proven/forced;
\textbf{T3} numerically verified, not theorem-level; \textbf{T4} explicitly falsified. The discipline's
non-negotiable rules: no claim is dressed in theorem language before its proof; negative results are
first-class outputs with the killing computation retained; every scope boundary is stated; overclaims are
corrected immediately. Two failure modes were guarded against throughout --- the \emph{sycophantic} mode (a
result that feels validated because a model agreed, not because it was tested) and the \emph{false-fit} mode
(a structural resemblance mistaken for an operational identity). Both were caught in practice and are
recorded.

\section{The representation: numbers as shapes}

\begin{definition}[Shape]
A \emph{shape} encodes a positive integer's multiplicative structure as a finite map $\{p_i:e_i\}$ from prime
sides to positive exponents, with volume $\prod_i p_i^{e_i}$. Multiplicative questions become geometric:
multiply $=$ merge (exponents add), divisibility $=$ containment, $\gcd/\mathrm{lcm}=$ common/enclosing box
(coordinatewise $\min/\max$), power $=$ scaling, primality $=$ atomic box. The only forced projection is
\texttt{volume()}; the only operation with no shape is addition, handled by an explicit flagged exit.
\end{definition}

\subsection{Reference engine (verified, complete source)}
The following is the entire reference implementation. All twelve operations are verified exact against
integer ground truth (Section~\ref{sec:ledger}). It transcribes directly to an integer-only Rust module.

\lstinputlisting{shape_engine_disp.py}

\section{The homomorphism criterion}

\begin{criterion}[Injective homomorphism into a pre-specified structure $\Rightarrow$ reversible]
Fix in advance a stored relational structure with an \emph{independently-meaningful} operation $\oplus$
(for the integers: exponent-vector addition, given by the Fundamental Theorem of Arithmetic before any
particular number is considered). \emph{(Forward, verified.)} If $\star$ is an \textbf{injective} homomorphism
$(\text{objects},\star)\to(\text{stored structure},\oplus)$ into that pre-specified structure --- a faithful
embedding, isomorphic onto its image --- then $\rho(a\star b)=\rho(a)\oplus\rho(b)$, the relational form
computes $\star$ without forming the scalar, and the map inverts: losslessly reversible. The qualifier
\textbf{injective} was forced by predictive testing (Section~\ref{sec:predictive}): a homomorphism alone
preserves the operation (density), a non-injective one loses identity (trace, determinant, entropy, a
barcode, the list-to-multiset map). \emph{(Converse, FALSE.)} Reversibility does \emph{not} imply injective
homomorphism: the Cantor pairing $\pi(a,b)=\tfrac{(a+b)(a+b+1)}{2}+b$ is a reversible bijection
$\mathbb{N}^2\!\to\!\mathbb{N}$ yet is not a homomorphism into the natural componentwise structure
($\pi(1,2)+\pi(2,0)=11\neq 17=\pi(3,2)$; adversarial counterexample, accepted). \emph{(Vacuity guard.)} The
target structure must be pre-specified: ``injective homomorphism into \emph{some} structure'' is empty, since
any bijection is an isomorphism onto the operation transported through it ($a\ast b:=\pi(\pi^{-1}a+\pi^{-1}b)$).
The criterion is therefore a one-way implication into a fixed, independently-meaningful target --- a
falsifiable empirical principle, not a biconditional.
\end{criterion}

For integers under multiplication this is the Fundamental Theorem of Arithmetic: factorization is an
isomorphism $(\mathbb{Z}_{>0},\times)\cong(\bigoplus_p\mathbb{N},+)$. Multiplication becomes exponent-vector
addition, $\gcd$ and $\mathrm{lcm}$ become $\min$ and $\max$, powers become scaling --- each verified
$\rho(ab)=\rho(a)\oplus\rho(b)$.

\begin{proposition}[Addition is non-homomorphic on factor structure, T1]
$\rho(a+b)$ is not a function of $\rho(a),\rho(b)$: $360+84=444$ and $363+81=444$ have identical sums from
disjoint summand structures, and a unit change ($360+86=446$) yields an unrelated factor structure.
Addition's flattening is therefore irreversible. This is the seam, stated as the exact non-homomorphic
boundary.
\end{proposition}

\section{Multiplication as folded addition; unique versus cardinal structure}

Multiplication on $\mathbb{N}$ is, by definition, iterated addition ($a\times(n{+}1)=a\times n + a$). The
content of the criterion is that \emph{folding addition to its irreducibles manufactures a unique structure
the unfolded operation lacks}: $12=2^2\cdot 3$ is a single canonical object, whereas $12$ as a sum is
realizable in many ways. Hence the two sides of the seam carry two different kinds of structure:
\begin{itemize}\setlength\itemsep{1pt}
\item \textbf{Unique} (multiplicative): the factorization is a point; FTA gives a canonical address.
\item \textbf{Cardinal} (additive): the decompositions form a cloud with no canonical address, but the
\emph{size} of the cloud is exact and deep --- the partition function $p(n)$ (e.g.\ $p(100)=190{,}569{,}292$).
\end{itemize}
Addition is not structureless; its structure is the count, not an address. Nor is the seam a total void
at the local level: the $p$-adic valuation obeys the ultrametric bound $v_p(a+b)\ge\min(v_p(a),v_p(b))$
(verified, $10^4$ cases), so the sum's shape is constrained to lie inside a lower-bound box at each prime.
Quantified (verified, $6\times10^4$ prime-coordinate cases): the bound is \emph{tight} --- $v_p(a+b)$ is
determined exactly --- whenever the two summands' valuations differ ($\approx 50\%$ of cases), and slack only
when they tie ($v_2(2+6)=3>\min=1$). The seam is therefore a \emph{gradient} of structural loss, not a binary
void: each prime-coordinate's valuation survives addition exactly off-ties and is lost only on-ties. Full
reversibility is still not restored (the on-tie cloud persists), but the irreversibility is partial and
measurable, not total.

\section{The seam characterized: convolution (keystone)}

\begin{theorem}[Convolution crosses the density side of the seam; components T1, identification T2/T3]
The additive (cardinal) structure is encoded multiplicatively by its generating function: Euler's identity
$\sum_n p(n)x^n=\prod_{k\ge 1}(1-x^k)^{-1}$ turns the entire additive cloud into one product (verified: the
product's coefficients equal $p(n)$ for $n\le 30$). The operation joining the two sides is \emph{convolution}:
multiplying generating functions convolves coefficients, $(a\,x^i)(b\,x^j)=ab\,x^{i+j}$, which \emph{adds} the
index and \emph{multiplies} the value (verified on coefficient sequences). The seam is therefore not
impassable: it is crossed by generating functions, and the crossing operation is convolution.
\end{theorem}

\begin{observation}[Density, not identity]
Convolution crosses the seam for counting and not for identifying: it yields how many, never which one. This
one property organizes three facts. The Riemann explicit formula is a convolution of primes against zeros,
so the prime/zero relationship is statistical (a $+122\sigma$ density signal) and never an address. Fully
homomorphic encryption \emph{illustrates} the seam (it does not prove it): both operations cannot be had by
gluing two single-operation schemes with incompatible ciphertext types, so a single rich structure
(lattice/LWE) must host both natively and pay in noise injected for lattice hardness (Section~\ref{sec:fhe}). And the additive wall is exactly the statement that the
sum's factor structure is the convolution of the summands' structures --- a distribution, not a point.
\end{observation}

\section{Predictive domain transfer: the criterion is a class, not a catalogue}\label{sec:predictive}

To test whether the criterion predicts reversibility \emph{outside} the examples it was built from, $30$
domains were chosen and their reversibility \emph{predicted in advance} (predictions locked before scoring),
then each was checked against ground truth. The set was deliberately stocked with operations that should
\emph{fail} (entropy, determinant, trace, hashing, PCA, neural embeddings, modular reduction, sorting) as
well as ones that should hold (Smith normal form, DFT, CRT, discrete log, logarithm, continued fractions,
run-length encoding, polynomial factorization).

\begin{observation}[$28/30$ predicted correctly; the misses sharpened the criterion]
Under the criterion as originally stated (``reversible $\leftarrow$ homomorphism''), $28$ of $30$ domains
were predicted correctly --- far above the $\sim\!50\%$ chance baseline, evidence that the criterion
describes a real class rather than retrospectively organizing chosen examples. The two misses, persistent
homology and the list-to-multiset map, were both predicted reversible and are in fact lossy, and they shared
a single flaw: each is a homomorphism that is \emph{not injective} (distinct filtrations share a barcode;
distinct lists share a multiset). Adding the injectivity qualifier (Criterion~3.1) corrects both, taking the
predictive score to $30/30$. The misses thus did not weaken the criterion; they forced its correct
statement.
\end{observation}

\begin{theorem}[Injectivity is the identity/density split, T2]
The corrected criterion unifies with the convolution seam. A homomorphism preserves the operation --- how
parts combine, the \emph{density} side. Injectivity is exactly the additional condition that distinct objects
have distinct images --- the \emph{identity} side. A non-injective homomorphism therefore carries density and
loses identity, which is precisely the seam's behaviour: convolution (a non-injective homomorphism of
generating functions) crosses for counts and not for which-one. The seam is thus not a special case but an
instance of the general mechanism: \emph{reversibility $=$ homomorphism (density) $+$ injectivity (identity)},
and the seam is where injectivity fails.
\end{theorem}

\section{Two independent axes: order-dependence and structure-retention}

A conjecture --- ``commutativity measures projected-away structure'' --- was tested and \emph{falsified} in
its strong form, yielding a sharper map. Non-commutativity measures retained \emph{order-dependent} structure:
scalar $+,\times$ are commutative, retaining none; the cross product is anti-commutative, retaining exactly
orientation (one bit, a structured sign flip); matrix products are fully non-commutative, retaining the whole
arrangement. But $\gcd$ and $\mathrm{lcm}$ are commutative \emph{and} retain full common multiplicative
structure. Hence order-dependence and structure-retention are \emph{independent axes}. The shape engine lives
in the commutative-and-structured cell, which is why its canonical fingerprint is path-independent
(commutativity) yet information-complete (structure).

\begin{center}
\begin{tabular}{lll}
\toprule
operation & order-dependent structure & order-independent structure \\
\midrule
scalar $+,\times$ & none & none \\
$\gcd,\mathrm{lcm}$, symmetric functions & none & full \\
cross product / determinant sign & one bit (orientation) & --- \\
matrix product, rotational path & full arrangement & --- \\
\bottomrule
\end{tabular}
\end{center}

\section{The verified ledger}\label{sec:ledger}

Every operation in the homomorphic class was computed in the relational form and checked exact against
standard-library ground truth (seed \texttt{20260423}); $14/14$ exact, $0$ failures.

\begin{center}
\begin{tabular}{lll}
\toprule
domain & operations & verification \\
\midrule
arithmetic functions & $d,\sigma,\sigma_2,\varphi,\mu$ & exact, $300$ values each \\
$\gcd$ / $\mathrm{lcm}$ & $\min$/$\max$ on exponents & exact, $300$ pairs each \\
binomial / factorial & Legendre exponent formula & $\binom{20}{7},\binom{100}{50},\binom{500}{123}$ exact \\
modular products & homomorphism under $\bmod$ & concentrated chain $\bmod\,(2^{61}{-}1)$, exact \\
lattice index / SNF & product of elementary divisors & $\sim\!1479$-digit index, exact, no overflow \\
multinomial multiplicity & $N!/\prod n_i!$ & $\sim\!4978$-digit $W$, exact \\
radical / perfect power & squarefree kernel, exponent $\gcd$ & exact, $200$ values \\
\midrule
seam keystone & Euler product $=p(n)$; convolution & exact, $n\le 30$; coefficient identity \\
\bottomrule
\end{tabular}
\end{center}

\noindent A clarifying correction (adversarial audit, second pass): the relational form answers
multiplicative questions about the multinomial $W$ ($\sim\!4978$ digits) and the $5000$-prime product
($\sim\!20975$ digits) \emph{without ever forming the integer}. We do \emph{not} claim the scalar is
incomputable: the integer builds in memory in milliseconds, and the only failure is at base-$10$ string
projection (Python's \texttt{CVE-2022-45061} default \texttt{int\_max\_str\_digits}); the shape engine's
own \texttt{volume()} hits the identical limit if asked to form the integer. The honest distinction is
therefore narrow and real: the relational workflow never \emph{needs} the projection step, while the scalar
workflow forms the integer as its normal representation. It is not a mathematical wall the relational form
transcends; it is a step the relational form does not require.

\subsection{The reproducible ledger (complete source)}
\lstinputlisting{verify_ledger_disp.py}

\subsection{The predictive test (complete source)}
The $30$-domain predictive transfer of Section~\ref{sec:predictive}, reproducible and self-checking
($28/30$ under the original criterion, $30/30$ under the injective correction).
\lstinputlisting{predictive_test_disp.py}

\section{The headline tractability instance: structure search beyond enumeration}

Searching exponent space for the integer $n\le N$ maximizing $d(n)$ never forms $n$. An exact
branch-and-bound with a \emph{provably admissible} bound (budget all remaining log-room on the cheapest
prime; each unit buys at most a factor of two) computes the true maximum.

\begin{center}
\begin{tabular}{lrr}
\toprule
$N$ & max $d(n)$ & time \\
\midrule
$10^{18}$ & $103{,}680$ & $7$\,ms \\
$10^{50}$ & $1.81\times10^{10}$ & $1.4$\,s \\
$10^{80}$ & $3.01\times10^{14}$ & $33$\,s \\
\bottomrule
\end{tabular}
\end{center}

\noindent Anchored exact against brute force at $N\le 60,840,720720,2162160$ ($d=12,32,240,320$). This reaches $N=10^{80}$
without forming $n$. A correction on attribution (adversarial audit): the tractability here is \emph{not} a
property of the relational representation. Since $d(n)=\prod_i(e_i+1)$ depends only on the exponents,
maximizing it under $\sum_i e_i\log p_i\le\log N$ is an unbounded-knapsack problem on exponent vectors that
any integer-array branch-and-bound solves equally fast; at $10^{80}$ the optimum has only $\sim\!54$ distinct
primes, so the search space is narrow. The result is a fact about highly composite numbers and an admissible
bound, not about shapes. We retain it as an example of computing a structural extremum without forming the
integer, while crediting the speed to the knapsack structure, not the representation. The path to this result is itself recorded:
a greedy heuristic (suboptimal, failed the anchor) and a non-admissible bound (pruned true optima, returned
wrong answers) were both falsified before the admissible version passed.

\section{The non-transitive class: when no scalar exists}

\begin{theorem}[No faithful scalar under cycles, T1]
If the pairwise majority relation contains a cycle $A\succ B\succ C\succ A$, every total order contradicts at
least one majority verdict (verified exhaustively: minimum contradicted edges is one, not zero). Total orders
are transitive; the relation is not; a transitive structure cannot realize a cycle.
\end{theorem}

This is the social-choice form of the obstruction that forbids encoding rock-paper-scissors as box
containment (a partial order cannot represent a cycle). Cycle frequency rises with options
($\sim\!7\%,43\%,78\%,99\%$ for $3,5,7,10$ candidates) and persists with electorate size ($\sim\!48\%$ at
$1001$ voters), so the impossibility is structural. The faithful representation is a directed dominance graph
--- integer, exact, supporting monotone elimination.

\subsection{The dominance engine and two-engine dispatcher (complete source)}
\lstinputlisting{dominance_engine_disp.py}

\section{Application analysis: homomorphic encryption}\label{sec:fhe}

The homomorphism of our criterion and the homomorphism of homomorphic encryption are the same algebraic
concept, and the connection is real but specific. Single-operation schemes are easy (RSA multiplies
ciphertexts; additive schemes add them, both verified on toy instances). Combining a multiplicative and an
additive scheme to obtain both fails: their ciphertext spaces are incompatible, and bridging them requires
either decryption (insecure) or a cheap structure-switch that does not exist --- the seam, relocated to the
tool boundary, not removed. The real solution (Gentry) uses one structure (lattice/LWE) native to both
operations and pays in noise management. The shape engine cannot be a ciphertext representation in any case:
it \emph{exposes} factorization, the exact opposite of what encryption must \emph{hide}. A correction on the
analogy (adversarial audit): mature FHE (Ring-LWE over $\mathbb{Z}_q[x]/(x^N+1)$) uses a \emph{single} rich
structure that natively supports both operations, and its dominant cost is multiplicative noise growth
injected for lattice-problem hardness --- not an add/multiply incompatibility. So the only defensible residue is a
known triviality: one cannot obtain both operations by \emph{gluing two single-operation schemes}, because
their ciphertext spaces are incompatible algebraic types --- a fact cryptographers never mistook for a
bridgeable gap. FHE is therefore an \emph{illustration} of the seam, not evidence for it. It is \emph{not}
that the seam explains FHE's expense; that expense
is noise management, a different mechanism. The earlier identification was an over-stretched analogy.

\section{The negative ledger (retained in full)}

\begin{center}
\begin{tabular}{lll}
\toprule
hypothesis & grade & how falsified \\
\midrule
multiply/divisibility faster than bigint & T4 & compiled C bigint wins to $\ge 39000$ bits \\
exact rationals beat \texttt{Fraction} & T4 & per-step C gcd beats dict merge ($11\times$) \\
near-unity ratios need exact shapes & T4 & log-subtraction is well-conditioned ($10^{-16}$) \\
greedy superabundant search & T4 & suboptimal; failed brute-force anchor \\
tighter (non-admissible) search bound & T4 & pruned true optima; returned wrong answers \\
BRA charge composes multiplicatively & T4 & kernel whitepaper: charge is additive table-sums \\
dual-engine geometric router & T4 & live source: ordinal thresholds, transitive relation \\
trust/qualia as multiplicative & T4 & additive aggregation (the seam) \\
Merkle composition as multiplicative & T4 & hashing destroys structure by design \\
``commutativity $=$ no structure'' & T4 & $\gcd$ is commutative and structured \\
 criterion converse (reversible $\Rightarrow$ inj.\ homo.) & T4 & Cantor pairing: reversible, not a homomorphism \\
 ``inj.\ homomorphism into \emph{some} structure'' & T4 & vacuous: any bijection via transported operation \\
$\varphi$+inverse-silver ``structures noise'' & T4 & $25\times$ worse than $\varphi$ alone (discrepancy) \\
FHT ``consciousness score'' detects structure & T4 & saturates to $1.0$ on noise and structure alike \\
\bottomrule
\end{tabular}
\end{center}

\noindent Four of these were architecture-integration hypotheses falsified against live OmniForge/AISO source
or the kernel specification; the clean structural resemblance was, in each case, the warning sign. Two were
early-work claims re-tested against ground truth and found to be a real kernel wrapped in unfalsifiable
naming. The final two are the criterion's own converse and a vacuity guard, both falsified by an independent
adversarial counterexample search (the Cantor pairing) and incorporated here rather than elsewhere. Each falsification is the criterion working: the homomorphism test, and the anchor, predict the
misses as well as the hits.

\section{Scope, limits, and open questions}

The mathematics invoked is classical (FTA, Legendre, multiplicativity of $\sigma,\varphi,\mu$, Smith normal
form, Euler's partition product, Arrow/Condorcet). The contribution is the criterion --- reversibility of a
flattening equals homomorphism into the stored relational structure --- its characterization of the seam as
convolution, and the verified record across seven applied domains, the impossibility class, and the wall.
\textbf{Limits, stated explicitly:} the principle is not an arithmetic accelerator (compiled big integers win
on throughput; the advantage is exactness, overflow avoidance, and faithful representation of the
unrepresentable); it requires an exact underlying relation (measured continuous data has none); its exactness
is worth its cost only where the question is itself exact or discrete; and the verified-optimal search reaches
$\sim\!10^{80}$, beyond which a tighter admissible bound is required and not yet constructed. The principle
does not touch primality or prime-counting (building a prime's shape is the factoring wall) and does not bear
on the Riemann Hypothesis (a categorical gap between finite multiplicative bookkeeping and the analytic
continuation of an infinite object). \textbf{Open:} a verified-optimal search past $10^{80}$; a
cleartext-side auditing role in multiplicatively-homomorphic protocols; and per-domain transposition tests in
fields not examined here, each to be run through the same gauntlet.

\section{Conclusion}

The thesis is a representational law with a checkable criterion: the algebraic flattening of a relational
object is losslessly reversible when the flattening is an injective homomorphism into a pre-specified structure (forward direction verified across the ledger and a $30$-domain predictive test; converse falsified by the Cantor pairing), and irreversible where injectivity fails (the seam),
and the seam is crossed for density by convolution while remaining uncrossable for identity. The result is offered
as a falsifiable program. Its verified instances, its impossibility class, its wall, and its fourteen retained
falsifications are all reproducible at seed \texttt{20260423}. What it asks of any reader is the same thing it
asked of its authors: do not believe the elegant statement; test it against ground truth, keep what survives,
and record what does not.

\vspace{1em}\noindent\rule{\linewidth}{0.4pt}\\
\small Consolidates \textit{The Geometry Engine}, \textit{When a Number Lies} (the relational principle and limits), and \textit{The Reversible Flattening} (the
homomorphism criterion and convolution seam). All source listings are the verified reference implementations.
SIP License v1.1. Seed \texttt{20260423}.
\end{document}
